---
title: "College Scorecard Analysis"
author: "Jaan Choudhri"
date: "5/16/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Codebook for College Scorecard Analysis


```{r cars}
library(dplyr)
library(ggplot2)
library(scales)
library(tidyverse)
```

```{r}
data <- read_csv("Most-Recent-Cohorts-All-Data-Elements.csv")
```
We have about 1800 schools with over 1700 variables.

We are only interested in a few categorical and quantitative variables. Here's how we can slim down our dataset to only focus on our desired variables:
  - award bachelor’s degrees
  - are not for-profit institutions
  - are currently operating.
  - are in the 50 states
```{r}
newData <- data %>%
  filter(PREDDEG == 3,
         CONTROL != 3,
         CURROPER == 1,
         ST_FIPS <= 56
         ) %>%
  select(school = INSTNM, institutionType = CONTROL, admit = ADM_RATE, med_income = MD_EARN_WNE_P10,
         med_fam_inc = MD_FAMINC, NPT4_PUB, NPT4_PRIV)
```
In order to properly evaluate our variables, we have to convert the quantitative variables to doubles instead of strings. 
```{r}
newData$NPT4_PUB <- as.double(newData$NPT4_PUB)
newData$NPT4_PRIV <- as.double(newData$NPT4_PRIV)
newData$med_income <- as.double(newData$med_income)
newData$med_fam_inc <- as.double(newData$med_fam_inc)
newData$admit <- as.double(newData$admit)
```
```{r}
newData <- newData %>%
  rowwise %>%
  mutate(cost = sum(NPT4_PUB, NPT4_PRIV, na.rm = TRUE)) %>%
  select(-c(NPT4_PUB, NPT4_PRIV))
newData["cost"][newData["cost"] == 0] <- NA
```
```{r}
head(newData)
```
A 1 for funding indicates a public school, while 2s indicate a private school. Admit is the rate of admission, med_earnings is the median income of a student following graduation, med_fam_inc is the students' family's median income, and price is the cost of attendance of the institution. 
```{r}
xdollar <- c(scale_x_continuous(labels = dollar,
                              breaks = seq(0, 130000, 25000),
                              limits = c(0, NA)))

ydollar <- c(scale_y_continuous(labels = dollar,
                                        breaks = seq(0, 130000, 25000),
                                        limits = c(0, NA)))

titling <- theme(plot.title = element_text(hjust = 0.5,
                                           face = "bold"),
                 axis.title.x = element_text(face = "bold"),
                 axis.title.y = element_text(face = "bold"))
```

```{r}
ggplot(data = newData) +
  geom_histogram(mapping = aes(x = med_income)) +
  labs(title = "Median Earnings of College Graduates",
       x = "Median Earnings (USD)",
       y = "Amount of Colleges") +
  xdollar +
  titling
```

```{r}
ggplot(data = newData,
       mapping = aes(x = med_fam_inc,
                     y = med_income)) +
  geom_point(size = 2,
             color = "skyblue2") +
  geom_smooth(method = "lm",
              color = "black") +
  labs(title = "Median Earnings of Graduates against \nMedian Family Income of Current Students",
       x = "Median Family Income in USD",
       y = "Median Earnings in USD") +
  xdollar +
  ydollar +
  titling
```

```{r}
point_theme <- c(scale_x_continuous(labels = percent),
                 ydollar,
                 scale_color_manual(labels = c("Public","Private"),
                                    values = c("#F8766D", "#00BFC4")))

scatter <- ggplot(data = newData,
                  mapping = aes(x = med_income,
                                y = med_fam_inc)) +
  geom_point(mapping = aes(color = factor(institutionType)),
             size = 1) +
  point_theme +
  titling
scatter +
  labs(title = "Median Earnings of Graduates against Admission Rate of Colleges \nSeparated by Source of Funding",
       x = "Median Earnings in USD",
       y = "Median Family Income",
       color = "Source of Funding") +
  facet_wrap(~ institutionType)
```

```{r}
fit2 <- lm(data = newData,
           med_income ~ med_fam_inc + institutionType + med_fam_inc*institutionType)
summary(fit2)
```